A Dependently Typed Multi-Stage Calculus

### 2019-08-28 text/related-work.tex
Relect all commments from Taha.

Theoretical studies on multi-stage programming owe a lot to seminal work by
Davies and Pfenning~\cite{DaviesPfenning01JACM} and
Davies~\cite{davies1996temporal}, who found Curry-Howard correspondence
between multi-stage calculi and modal logic. In particular, Davies'
$\lambda^\circ$~\cite{davies1996temporal} has been a basis for several
multi-stage calculi with quasi-quotation. $\lambda^\circ$ did not have
operators for run and CSP; a few
studies~\cite{benaissa1999logical,MoggiTahaBenaissaSheard99ESOP} enhanced and
improved $\lambda^\circ$ towards the development of a type-safe multi-stage
calculus with quasi-quotation, run, and CSP, which were proposed by Taha and
Sheard as constructs for multi-stage programming~\cite{MetaML}.

@@

Although there are studies on combinations of multi-stage programming and
other programming features such as mutable cells~\cite{kiselyov2016refined} and
control operators~\cite{KameyamaKiselyovShan09PEPM,oishi2017staging}, a
combination with dependent types has been little studied.

Brady and Hammond~\cite{brady2006dependently} have discussed a
combination of dependently typed programming with staging in the style of
MetaOCaml to implement a staged interpreter, which is statically guaranteed
to generate well-typed code. However, they focused on concrete programming
examples and there is no theoretical investigation of the programming
language they used.

Pasalic and Taha~\cite{pasalic2002tagless} designed \(\lambda_{H\circ}\) by
introducing the concept of stage into an existing dependent type system
\(\lambda_H\)~\cite{zhong2002certified}. However, there are not eval and CSP
in \(\lambda_{H\circ}\) because it was designed before the invention of
environment classifiers.

In Forgarty, Pasalic, Siek and Taha~\cite{fogarty2007concoqtion}

@@

Although we define type equivalence of \LMD with composition of equivalence
rules, there is another candidate to define, which gives reduction on types
and compare the results of reduction such as~\cite{sorensen2006lectures}.
This method is better than one of \LMD because equivalence rules become
simple. However, we reject it because it cannot handle CSP flexibly.

>>>>

Theoretical studies on multi-stage programming owe a lot to seminal work by
Davies and Pfenning~\cite{DaviesPfenning01JACM} and
Davies~\cite{davies1996temporal}, who found Curry-Howard correspondence
between multi-stage calculi and modal logic. In particular, Davies'
$\lambda^\circ$~\cite{davies1996temporal} has been a basis for several
multi-stage calculi with quasi-quotation. $\lambda^\circ$ did not have
operators for run and CSP; a few
studies~\cite{benaissa1999logical,MoggiTahaBenaissaSheard99ESOP} enhanced and
improved $\lambda^\circ$ towards the development of a type-safe multi-stage
calculus with quasi-quotation, run, and CSP, which were proposed by Taha and
Sheard as constructs for multi-stage programming~\cite{MetaML}. Chen and
Xi~\cite{chen2003meta} introduce a code type with types of free variables in
the code in order to prevent free variable evaluation. \AK{Is here
appropriate? Although this paper using the concept of dependent types, the
purpose is avoid open code problem. Therefore, I cite this paper here.}

@@

Although there are studies on combinations of multi-stage programming and
other programming features such as mutable cells~\cite{kiselyov2016refined}
and control operators~\cite{KameyamaKiselyovShan09PEPM,oishi2017staging}, a
combination with dependent types has been little studied. \AK{little?} Brady
and Hammond~\cite{brady2006dependently} have discussed a combination of
dependently typed programming with staging in the style of MetaOCaml to
implement a staged interpreter, which is statically guaranteed to generate
well-typed code. However, they focused on concrete programming examples and
there is no theoretical investigation of the programming language they used.
Pasalic and Taha~\cite{pasalic2002tagless} designed \(\lambda_{H\circ}\) by
introducing the concept of stage into an existing dependent type system
\(\lambda_H\)~\cite{zhong2002certified}. However, there are not eval and CSP
in \(\lambda_{H\circ}\) because it was designed before the invention of
environment classifiers. In Forgarty, Pasalic, Siek and
Taha~\cite{fogarty2007concoqtion}, they extended the type system of MetaOCaml
with indexed types. With this extension, types can be indexed with a Coq term.

@@

% Although we define type equivalence of \LMD with composition of equivalence
% rules, there is another candidate to define, which gives reduction on types
% and compare the results of reduction such as~\cite{sorensen2006lectures}.
% This method is better than one of \LMD because equivalence rules become
% simple. However, we reject it because it cannot handle CSP flexibly.
